Follow-up Samples

### 2017-01-18 submission/manuscript_outline_20161024.Rmd
streamline -results

***Bacterial Community and FIT Changes before and after Surgery***
Based on thetayc distance metrics, comparing the initial to the follow up samples, there was a significant difference between the adenoma and carcinoma groups (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`) [Figure 1a]. There was also a significant difference in change in FIT between initial and follow up samples (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`) [Figure 1b].  The whole community structure before and after surgery were visualized on NMDS graphs for both adenoma [Figure1c] (PERMANOVA = `r format(beta_summary['Adenoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) and carcinoma [Figure 1d] (PERMANOVA = `r format(beta_summary['Carcinoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  When all initial and follow up samples were compared to each other there was no significant overall difference between them (PERMANOVA = `r format(beta_summary['All', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  There was no significant difference between initial and follow up samples for observed OTUs, Shannon diversity, or evenness after correction for multiple comparisons [Table S1].  There was also no significant difference between initial and follow up samples for any single OTU [Figure S1].

***Differences in Adenoma and Carcinoma of Previously Associated Cancer Bacteria***
We next examined whether there were differences in previously well described carcinoma associated OTUs.  These included the OTUs that aligned with *Porphyromonas asaccharolytica* (`r unique(filter(crc_bug_info, tax_id == "Porphyromonas")[, "otu"])`), *Fusobacterium nucleatum* (`r unique(filter(crc_bug_info, tax_id == "Fusobacterium")[, "otu"])`), *Parvimonas micra* (`r unique(filter(crc_bug_info, tax_id == "Parvimonas")[, "otu"])`), and *Peptostreptococcus stomatis* (`r unique(filter(crc_bug_info, tax_id == "Peptostreptococcus")[, "otu"])`).  First, the carcinoma samples showed a significant difference between initial and follow up samples for *Peptostreptococcus stomatis* (P-value = `r format(main_crc_bugs_summary["pept", "crc_BH"], digits = 3)`) and *Porphyromonas asaccharolytica* (P-value = `r format(main_crc_bugs_summary["porp", "crc_BH"], digits = 3)`) whereas there was no signficant differences in any of these OTUs in the adenoma samples [Table S4]. Second, when these OTUs were present, there was a clear magnitude difference based on whether they were from adenoma or carcinoma [Figure 2].  However, only a small percentage of those with adenoma or carcinoma were positive for any of these OTUs.

***Full and Reduced Model Results***
Since differences were observed between initial and follow up samples and only a small number of individuals were positive for previously associated CRC bacteria; we next investigated if we could create a model that could adequately classify and adjust lesion probability based on the bacterial community and FIT.  The lesion model had an AUC range of `r format(min(lesion_AUC_run_summary$ROC), digits = 3)` to `r format(max(lesion_AUC_run_summary$ROC), digits = 3)` versus the initial follow up model which had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 20 repeated 10-fold cross validations.  Interestingly, identification of the most important variables and reducing the models to only these factors increased the AUC in the lesion model (`r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) and initial and follow up model (`r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)`).

The test set AUC range for the full and reduced lesion model were similar to that reported for the training set AUC ranges and the ROC curve ranges overlap each other [Figure 3a].  The ROC curve for the final lesion model used was within the range of both the full and reduced lesion model [Figure 3a].  Interestingly, the test set AUC range for the initial and follow up performed much better then the training set AUCs.  Both the full and reduced initial and follow up models overlapped with each other [Figure 3b] there was a marked decrease in the ROC curve for the final model used.

***Most Important Variables to the Models***
The reduced models were built based on the most important variables to the respective full model.  For the lesion model there were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2] whereas for the initial and follow up model there were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S3].  For both models FIT resulted in the largest decrease in MDA [Figure S2a & S3a].

***Positive Probability Prediction after Surgical Removal of Adenoma or Carcinoma***
Regardless of model used there was a significant decrease in the positive probability of either the sample being lesion or an initial sample on follow up [Figure 4 & S4] (full lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, reduced lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, and reduced initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).

For the full and reduced lesion model there was a significant difference in the classification for the lesion model between predicted and actual (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)`, respectively) but not for the initial follow up model (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)`).  However, the lesion model correctly kept the one individual who still had a carcinoma on follow up above the cut off point [Figure 4a & S4a] for a positive call while the initial and follow up models did not [Figure 4b & S4b].

***Common OTUs to both Lesion and Initial and Follow Up Models***
There were a total of `r length(common_vars_summary$otu)` OTUs that were common to both models.  Of these OTUs the most common taxonomic identifications were to  `r common_vars_summary['Otu000012', 'lowest_ID']`, `r common_vars_summary['Otu000005', 'lowest_ID']`, `r common_vars_summary['Otu000082', 'lowest_ID']`, and `r paste(gsub("_unclassified", "", common_vars_summary['Otu000006', 'lowest_ID']))`.  The majority of these OTUs had taxonomic identification to bacteria typically thought of as commensal [Table S2].

***Treatment and Time Differences***
There was no difference in the amount of change in positive probability for either the full or reduced lesion model for either chemotherapy (P-value = `r format(chemo_rad_summary['chemo_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['chemo_red_les', 'bh'], digits = 3)`) or radiation therapy (P-value = `r format(chemo_rad_summary['rad_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['rad_red_les', 'bh'], digits = 3)`). Although the initial follow up model was similar there was a significant decrease in positive probability for those treated with chemotherapy (P-value = `r format(chemo_rad_summary['chemo_IF', 'bh'], scientific = TRUE, digits = 3)` and `r format(chemo_rad_summary['chemo_red_IF', 'bh'], scientific = TRUE, digits = 3)`).  Time elapsed between the collection of the follow up sample from the initial sample, did not have a significant difference between adenoma and carcinoma (uncorrected P-value = `r format(time_pvalues["thetayc_change", "uncorrected_p_value"], digits = 3)`).

>>>>

***Bacterial Community and FIT***
Based on the thetayc distance metric, comparing the initial to the follow up samples, there was a significant difference between the adenoma and carcinoma groups (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`) [Figure 1a]. There was also a significant difference in change in FIT between initial and follow up samples (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`) [Figure 1b].  The whole community structure before and after surgery was visualized on NMDS graphs for both adenoma [Figure1c] (PERMANOVA = `r format(beta_summary['Adenoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) and carcinoma [Figure 1d] (PERMANOVA = `r format(beta_summary['Carcinoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  When all initial and follow up samples were compared to each other there was no significant overall difference between them (PERMANOVA = `r format(beta_summary['All', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  There was no significant difference between initial and follow up samples for observed OTUs, Shannon diversity, or evenness after correction for multiple comparisons [Table S1].  There was also no significant difference between initial and follow up samples for any single OTU [Figure S1].

***Previously Associated Cancer Bacteria***
We next examined whether there were differences in previously well described carcinoma associated OTUs.  These included the OTUs that aligned with *Porphyromonas asaccharolytica* (`r unique(filter(crc_bug_info, tax_id == "Porphyromonas")[, "otu"])`), *Fusobacterium nucleatum* (`r unique(filter(crc_bug_info, tax_id == "Fusobacterium")[, "otu"])`), *Parvimonas micra* (`r unique(filter(crc_bug_info, tax_id == "Parvimonas")[, "otu"])`), and *Peptostreptococcus stomatis* (`r unique(filter(crc_bug_info, tax_id == "Peptostreptococcus")[, "otu"])`).  First, the carcinoma samples showed a significant difference between initial and follow up samples for *Peptostreptococcus stomatis* (P-value = `r format(main_crc_bugs_summary["pept", "crc_BH"], digits = 3)`) and *Porphyromonas asaccharolytica* (P-value = `r format(main_crc_bugs_summary["porp", "crc_BH"], digits = 3)`) whereas there were no signficant differences in any of these OTUs in the adenoma samples [Table S4]. Second, when these OTUs were present, there was a clear magnitude difference based on whether they were from adenoma or carcinoma [Figure 2].  However, only a small percentage of those with adenoma or carcinoma were positive for any of these OTUs.

***Full and Reduced Model***
Since differences were observed between initial and follow up samples and only a small number of individuals were positive for previously associated CRC bacteria; we next investigated if we could create models that could adequately classify and adjust either lesion or before sample probability based on the bacterial community and FIT.  The lesion model had an AUC range of `r format(min(lesion_AUC_run_summary$ROC), digits = 3)` to `r format(max(lesion_AUC_run_summary$ROC), digits = 3)` versus the before sample model which had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  Interestingly, identification of the most important variables and reducing the models to only these factors increased the AUC in the lesion model (`r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) and before sample model (`r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)`).

The test set AUC range for the full and reduced lesion model were similar to that reported for the training set AUC ranges and the ROC curve ranges overlap each other [Figure 3a].  The ROC curve for the final lesion model used was within the range of both the full and reduced lesion model [Figure 3a].  Interestingly, the test set AUC range for the before sample model performed much better then the training set AUCs.  Both the full and reduced before sample models overlapped with each other [Figure 3b] there was a marked decrease in the ROC curve for the final model used.

***Most Important Variables***
The reduced models were built based on the most important variables to the respective full model.  For the lesion model there were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2] whereas for the before sample model there were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S3].  For both models FIT resulted in the largest decrease in MDA [Figure S2a & S3a].

***Positive Probability after Lesion Removal***
Regardless of model used there was a significant decrease in the positive probability of either the sample being lesion or a before sample on follow up [Figure 4 & S4] (full lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, reduced lesion P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`, and reduced initial and follow up P-value = `r format(filter(all_probs_paired_wilcox, model_type =="red_IF" & comparison == "lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).

For the full and reduced lesion model there was a significant difference in the classification for the lesion model between predicted and actual (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"], scientific = TRUE, digits = 3)`, respectively) but not for the before sample model (P-value = `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)` and `r format(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_IF" & info == "McnemarPValue")[, "all"], digits = 3, nsmall = 2)`).  However, the lesion model correctly kept the one individual who still had a carcinoma on follow up above the cut off threshold [Figure 4a & S4a] for a positive call while the before sample models did not [Figure 4b & S4b].

***Common OTUs to both Models***
There were a total of `r length(common_vars_summary$otu)` OTUs that were common to both models.  Of these OTUs the most common taxonomic identifications were to  `r common_vars_summary['Otu000012', 'lowest_ID']`, `r common_vars_summary['Otu000005', 'lowest_ID']`, `r common_vars_summary['Otu000082', 'lowest_ID']`, and `r paste(gsub("_unclassified", "", common_vars_summary['Otu000006', 'lowest_ID']))`.  The majority of these OTUs had classifications to bacteria typically thought of as commensal [Table S2].

***Treatment and Time Differences***
There was no difference in the amount of change in positive probability for either the full or reduced lesion model for either chemotherapy (P-value = `r format(chemo_rad_summary['chemo_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['chemo_red_les', 'bh'], digits = 3)`) or radiation therapy (P-value = `r format(chemo_rad_summary['rad_les', 'bh'], digits = 3)` and `r format(chemo_rad_summary['rad_red_les', 'bh'], digits = 3)`). Although the before sample model was similar there was a significant decrease in positive probability for those treated with chemotherapy (P-value = `r format(chemo_rad_summary['chemo_IF', 'bh'], scientific = TRUE, digits = 3)` and `r format(chemo_rad_summary['chemo_red_IF', 'bh'], scientific = TRUE, digits = 3)`).  Time elapsed between the collection of the follow up sample from the initial sample, did not have a significant difference between adenoma and carcinoma (uncorrected P-value = `r format(time_pvalues["thetayc_change", "uncorrected_p_value"], digits = 3)`).

### 2017-01-18 submission/manuscript_outline_20161024.tex
streamline -results

\textbf{\emph{Bacterial Community and FIT Changes before and after
Surgery}} Based on thetayc distance metrics, comparing the initial to
the follow up samples, there was a significant difference between the
adenoma and carcinoma groups (P-value = 0.00198) {[}Figure 1a{]}. There
was also a significant difference in change in FIT between initial and
follow up samples (P-value = 2.15e-05) {[}Figure 1b{]}. The whole
community structure before and after surgery were visualized on NMDS
graphs for both adenoma {[}Figure1c{]} (PERMANOVA = 0.997) and carcinoma
{[}Figure 1d{]} (PERMANOVA = 0.002). When all initial and follow up
samples were compared to each other there was no significant overall
difference between them (PERMANOVA = 0.085). There was no significant
difference between initial and follow up samples for observed OTUs,
Shannon diversity, or evenness after correction for multiple comparisons
{[}Table S1{]}. There was also no significant difference between initial
and follow up samples for any single OTU {[}Figure S1{]}.

\textbf{\emph{Differences in Adenoma and Carcinoma of Previously
Associated Cancer Bacteria}} We next examined whether there were
differences in previously well described carcinoma associated OTUs.
These included the OTUs that aligned with \emph{Porphyromonas
asaccharolytica} (Otu000153), \emph{Fusobacterium nucleatum}
(Otu000226), \emph{Parvimonas micra} (Otu000460), and
\emph{Peptostreptococcus stomatis} (Otu000653). First, the carcinoma
samples showed a significant difference between initial and follow up
samples for \emph{Peptostreptococcus stomatis} (P-value = 0.0183) and
\emph{Porphyromonas asaccharolytica} (P-value = 0.0154) whereas there
was no signficant differences in any of these OTUs in the adenoma
samples {[}Table S4{]}. Second, when these OTUs were present, there was
a clear magnitude difference based on whether they were from adenoma or
carcinoma {[}Figure 2{]}. However, only a small percentage of those with
adenoma or carcinoma were positive for any of these OTUs.

\textbf{\emph{Full and Reduced Model Results}} Since differences were
observed between initial and follow up samples and only a small number
of individuals were positive for previously associated CRC bacteria; we
next investigated if we could create a model that could adequately
classify and adjust lesion probability based on the bacterial community
and FIT. The lesion model had an AUC range of 0.723 to 0.795 versus the
initial follow up model which had an AUC range of 0.451 to 0.67 after
100 20 repeated 10-fold cross validations. Interestingly, identification
of the most important variables and reducing the models to only these
factors increased the AUC in the lesion model (0.791 - 0.847) and
initial and follow up model (0.651 - 0.79).

The test set AUC range for the full and reduced lesion model were
similar to that reported for the training set AUC ranges and the ROC
curve ranges overlap each other {[}Figure 3a{]}. The ROC curve for the
final lesion model used was within the range of both the full and
reduced lesion model {[}Figure 3a{]}. Interestingly, the test set AUC
range for the initial and follow up performed much better then the
training set AUCs. Both the full and reduced initial and follow up
models overlapped with each other {[}Figure 3b{]} there was a marked
decrease in the ROC curve for the final model used.

\textbf{\emph{Most Important Variables to the Models}} The reduced
models were built based on the most important variables to the
respective full model. For the lesion model there were a total of 37
variables {[}Figure S2{]} whereas for the initial and follow up model
there were a total of 33 variables {[}Figure S3{]}. For both models FIT
resulted in the largest decrease in MDA {[}Figure S2a \& S3a{]}.

\textbf{\emph{Positive Probability Prediction after Surgical Removal of
Adenoma or Carcinoma}} Regardless of model used there was a significant
decrease in the positive probability of either the sample being lesion
or an initial sample on follow up {[}Figure 4 \& S4{]} (full lesion
P-value = 1.11e-11, reduced lesion P-value = 1.91e-11, initial and
follow up P-value = 6.71e-12, and reduced initial and follow up P-value
= 6.72e-12).

For the full and reduced lesion model there was a significant difference
in the classification for the lesion model between predicted and actual
(P-value = 4.19e-10 and 6.98e-10, respectively) but not for the initial
follow up model (P-value = 1.00 and 1.00). However, the lesion model
correctly kept the one individual who still had a carcinoma on follow up
above the cut off point {[}Figure 4a \& S4a{]} for a positive call while
the initial and follow up models did not {[}Figure 4b \& S4b{]}.

\textbf{\emph{Common OTUs to both Lesion and Initial and Follow Up
Models}} There were a total of 14 OTUs that were common to both models.
Of these OTUs the most common taxonomic identifications were to Blautia,
Bacteroides, Streptococcus, and Clostridiales. The majority of these
OTUs had taxonomic identification to bacteria typically thought of as
commensal {[}Table S2{]}.

\textbf{\emph{Treatment and Time Differences}} There was no difference
in the amount of change in positive probability for either the full or
reduced lesion model for either chemotherapy (P-value = 0.821 and 0.821)
or radiation therapy (P-value = 0.69 and 0.981). Although the initial
follow up model was similar there was a significant decrease in positive
probability for those treated with chemotherapy (P-value = 7.04e-04 and
5.07e-03). Time elapsed between the collection of the follow up sample
from the initial sample, did not have a significant difference between
adenoma and carcinoma (uncorrected P-value = 0.784).

>>>>

\textbf{\emph{Bacterial Community and FIT}} Based on the thetayc
distance metric, comparing the initial to the follow up samples, there
was a significant difference between the adenoma and carcinoma groups
(P-value = 0.00198) {[}Figure 1a{]}. There was also a significant
difference in change in FIT between initial and follow up samples
(P-value = 2.15e-05) {[}Figure 1b{]}. The whole community structure
before and after surgery was visualized on NMDS graphs for both adenoma
{[}Figure1c{]} (PERMANOVA = 0.997) and carcinoma {[}Figure 1d{]}
(PERMANOVA = 0.002). When all initial and follow up samples were
compared to each other there was no significant overall difference
between them (PERMANOVA = 0.085). There was no significant difference
between initial and follow up samples for observed OTUs, Shannon
diversity, or evenness after correction for multiple comparisons
{[}Table S1{]}. There was also no significant difference between initial
and follow up samples for any single OTU {[}Figure S1{]}.

\textbf{\emph{Previously Associated Cancer Bacteria}} We next examined
whether there were differences in previously well described carcinoma
associated OTUs. These included the OTUs that aligned with
\emph{Porphyromonas asaccharolytica} (Otu000153), \emph{Fusobacterium
nucleatum} (Otu000226), \emph{Parvimonas micra} (Otu000460), and
\emph{Peptostreptococcus stomatis} (Otu000653). First, the carcinoma
samples showed a significant difference between initial and follow up
samples for \emph{Peptostreptococcus stomatis} (P-value = 0.0183) and
\emph{Porphyromonas asaccharolytica} (P-value = 0.0154) whereas there
were no signficant differences in any of these OTUs in the adenoma
samples {[}Table S4{]}. Second, when these OTUs were present, there was
a clear magnitude difference based on whether they were from adenoma or
carcinoma {[}Figure 2{]}. However, only a small percentage of those with
adenoma or carcinoma were positive for any of these OTUs.

\textbf{\emph{Full and Reduced Model}} Since differences were observed
between initial and follow up samples and only a small number of
individuals were positive for previously associated CRC bacteria; we
next investigated if we could create models that could adequately
classify and adjust either lesion or before sample probability based on
the bacterial community and FIT. The lesion model had an AUC range of
0.723 to 0.795 versus the before sample model which had an AUC range of
0.451 to 0.67 after 100 iterations of 20 repeated 10-fold cross
validations. Interestingly, identification of the most important
variables and reducing the models to only these factors increased the
AUC in the lesion model (0.791 - 0.847) and before sample model (0.651 -
0.79).

The test set AUC range for the full and reduced lesion model were
similar to that reported for the training set AUC ranges and the ROC
curve ranges overlap each other {[}Figure 3a{]}. The ROC curve for the
final lesion model used was within the range of both the full and
reduced lesion model {[}Figure 3a{]}. Interestingly, the test set AUC
range for the before sample model performed much better then the
training set AUCs. Both the full and reduced before sample models
overlapped with each other {[}Figure 3b{]} there was a marked decrease
in the ROC curve for the final model used.

\textbf{\emph{Most Important Variables}} The reduced models were built
based on the most important variables to the respective full model. For
the lesion model there were a total of 37 variables {[}Figure S2{]}
whereas for the before sample model there were a total of 33 variables
{[}Figure S3{]}. For both models FIT resulted in the largest decrease in
MDA {[}Figure S2a \& S3a{]}.

\textbf{\emph{Positive Probability after Lesion Removal}} Regardless of
model used there was a significant decrease in the positive probability
of either the sample being lesion or a before sample on follow up
{[}Figure 4 \& S4{]} (full lesion P-value = 1.11e-11, reduced lesion
P-value = 1.91e-11, initial and follow up P-value = 6.71e-12, and
reduced initial and follow up P-value = 6.72e-12).

For the full and reduced lesion model there was a significant difference
in the classification for the lesion model between predicted and actual
(P-value = 4.19e-10 and 6.98e-10, respectively) but not for the before
sample model (P-value = 1.00 and 1.00). However, the lesion model
correctly kept the one individual who still had a carcinoma on follow up
above the cut off threshold {[}Figure 4a \& S4a{]} for a positive call
while the before sample models did not {[}Figure 4b \& S4b{]}.

\textbf{\emph{Common OTUs to both Models}} There were a total of 14 OTUs
that were common to both models. Of these OTUs the most common taxonomic
identifications were to Blautia, Bacteroides, Streptococcus, and
Clostridiales. The majority of these OTUs had classifications to
bacteria typically thought of as commensal {[}Table S2{]}.

\textbf{\emph{Treatment and Time Differences}} There was no difference
in the amount of change in positive probability for either the full or
reduced lesion model for either chemotherapy (P-value = 0.821 and 0.821)
or radiation therapy (P-value = 0.69 and 0.981). Although the before
sample model was similar there was a significant decrease in positive
probability for those treated with chemotherapy (P-value = 7.04e-04 and
5.07e-03). Time elapsed between the collection of the follow up sample
from the initial sample, did not have a significant difference between
adenoma and carcinoma (uncorrected P-value = 0.784).
